Sequence Divergence as the main modulator of structure-sequence correlations

### 2014-10-30 manuscript/cordiv.tex
some new analysis and data

    We calculate the relevant Voronoi cell properties of all sites in all proteins using the software VORO++ (xx), and use DSSP (xx) for the calculation of Accessible Surface Area (ASA) for each site normalized by the theoretical maximum solvent accessibility values of Tein et al (20112 xx) to obtain the Relative Solvent Accessibility (RSA) for all individual sites in all proteins. In addition to ASA values, we also extract from DSSP information about the secondary structure of proteins, such as the total number of residues participating in different types of helices, parallel or anti-parallel beta sheets, or loops and turns. To complete the list of pdb-level structural properties, we also calculate the Spearman correlations between all residue-level structural properties and include them in the analysis to probe their potential effects on the strength of structure-sequence relations.
    \\

@@

        For the measure of local packing density in proteins (the Weighted Contact Number) we find that among all possible set of coordinates, the average over coordinates of all heavy atoms for each each individual side chain results in WCN values that show the best correlation with other structural and sequence properties, such as RSA, Voronoi cell properties, sequence entropy, and evolutionary rates. Specifically, WCN from average side chain coordinates (wcnSC) outperforms WCN based on $C_\alpha$ coordinates (wcnCA) in predicting RSA, Voronoi cell volumes, sequence entropy and evolutionary rates by a median Spearman correlation difference of $00$, $00$, $00$ \& $00$, respectively (Figure \ref{fig:best_wcn}).
        \\

        For the measure of local flexibility in proteins (Bfactor) we similarly find that among all $7$ representative measures of site Bfactors, the average of Bfactor values over all heavy atoms of each individual side chain (bfSC) results in the best correlations with other structural and sequence properties. Specifically, bfSC outperforms the commonly used $C_\alpha$ Bfactor (bfCA) in predicting RSA, Voronoi cell volumes, sequence entropy and evolutionary rates by a median Spearman correlation difference of $00$, $00$, $00$ \& $00$, respectively (Figure \ref{fig:best_bf}).
        \\

@@

        Based on the observation that side chain structural properties have the best correlations with other structure and sequence characteristics of individual sites, we keep only variables measured based on average side chain properties and coordinates throughout the rest of the analysis and omit all other possible measures that show only weaker correlations with other site-specific characteristics.
        \\

@@

>>>>

    We use VORO++ software (xx) to calculate the relevant Voronoi cell properties of all sites in all proteins, and use DSSP (xx) for the calculation of Accessible Surface Area (ASA) for each site normalized by the theoretical maximum solvent accessibility values of Tein et al (20112 xx) to obtain the Relative Solvent Accessibility (RSA) for all individual sites in all proteins. In addition to ASA values, we also extract from DSSP output, information about the secondary structure of proteins such as the total number of residues participating in different types of helices, parallel or anti-parallel beta sheets, or loops and turns. To complete the list of pdb-level structural properties, we also calculate the Spearman correlations between all residue-level structure and sequence properties and include them in the analysis to probe their potential effects on the strength of structure-sequence relations.
    \\

@@

        For the measure of local packing density in proteins (the Weighted Contact Number) we find that among all possible set of coordinates, the average over coordinates of all heavy atoms of each individual side chain results in WCN values that show the best correlation with other structural and sequence properties, such as RSA, Voronoi cell properties, sequence entropy, and evolutionary rates. Specifically, WCN from average side chain coordinates (wcnSC) outperforms WCN based on $C_\alpha$ coordinates (wcnCA) in predicting RSA, $\Delta\Delta$ entropy, sequence entropy and evolutionary rates ($r4sJC$) by a median Spearman correlation difference of $0.09$, $0.10$, $0.07$ \& $0.08$, respectively (Figure \ref{fig:best_wcn}).
        \\

        For the measure of local flexibility in proteins (Bfactor) we similarly find that among all $7$ representative measures of site Bfactors, the average of Bfactor values over all heavy atoms of each individual side chain (bfSC) results in the best correlations with other structural and sequence properties. Specifically, bfSC outperforms the commonly used $C_\alpha$ Bfactor (bfCA) in predicting RSA, $\Delta\Delta$ entropy, sequence entropy and evolutionary rates by a median Spearman correlation difference of $0.11$, $0.12$, $0.08$ \& $0.09$, respectively (Figure \ref{fig:best_bf}).
        \\

@@

        Based on the observations described in the previous paragraphs, we keep only variables measured based on average side chain properties and coordinates throughout the rest of the analysis and omit all other possible measures that show only weaker correlations with other site-specific characteristics.
        \\

@@

        The exclusion of alternative measures of structural quantities as described in the previous section, results in a significant reduction in the number of pdb-level variables to be further analysed, from $$ to $$, without compromising generality and comprehensiveness of the analysis.

        Now with the reduced data set at hand, we attempt to identify the main determinants of each sequence-structure relation of interest using regularized regression methods (xx) on the original dataset of pdb-level properties. , also on its rank transformation to ensure non-linearities

        first remove all other sequence-structure correlations from the list of potential contributing factors. As explained in the previous sections, Then we apply regularized regression methods
